---
title: "Week 3 Assignment"
subtitle: "Life Expectancy: Best, Worst, and Most Improved"
author: "Fenton Taylor"
date: "December 22, 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(dplyr)
library(reshape2)
```

## The Data
The data for this analysis is from: [data.worldbank.org](http://data.worldbank.org/indicator/SP.DYN.LE00.IN).

- The dataset contains the average life expectancy for 264 countries and regions from 1960-2014. Some of the countries have incomplete data, and will be excluded from the analysis.
- `meanLE`, the average of life expectancy from 1960-2014, and `changeInLE`, the difference in life expectancy from 1960-2014, have been calculated and added to the data.

### A small sample of the data:

```{r, message=FALSE, warning=FALSE, comment=""}
life <- read.csv("lifeExp.csv")
life <- life[,-(60:61)]
life <- life[complete.cases(life)==TRUE,]
life$meanLE <- rowMeans(life[,5:59])
life$changeInLE <- life$X2014-life$X1960
life[life$Country.Name%in%c("Aruba","Brazil","Canada","Denmark","Ethiopia","Finland","Germany", "Saudi Arabia", "Zimbabwe"), c(1,5:6,59:61)]
```

## Best and Worst
```{r}
life <- arrange(life, desc(meanLE))
bw <- life[c(1:10,(nrow(life)-9):nrow(life)),c(1,2,60)]
bw$class <- gl(2,10, labels=c("Best 10", "Worst 10"))
bw$Country.Name <- factor(bw$Country.Name, levels=bw$Country.Name[order(bw$meanLE)])
g <- ggplot(bw, aes(x=Country.Name, y=meanLE, fill=class)) + 
    theme_minimal()+
    geom_bar(stat = "identity")+
    scale_fill_manual(values=c("green3", "firebrick2"))+
    coord_flip() +
    labs(y='Mean Life Expectancy', x='Country', title="Ten Best and Worst Countries\nFor Life Expectancy (1960-2014)") +
    theme(legend.position = 'none')
p <- ggplotly(g)
p
```

## Most Improved
```{r}
mi <- arrange(life, desc(changeInLE))[1:12,-c(2:4,60:61)]
miMelt <- melt(mi, id.vars=1, measure.vars=2:56, value.name='LifeExp')
miMelt$Year <- rep(1960:2014, each=12)
g2 <- ggplot(miMelt, aes(x=Year, y=LifeExp,color=Country.Name))+
    geom_line()+
    labs(y='Life Expectancy', title='Top 12 Countries with Most Improved Life Expectancy')+
    theme(legend.title = element_blank())
p2 <- ggplotly(g2)
p2

```

## The End
Thanks for checking it out. Here is the link to the github repo, in case you are 
interested in checking out the code for the presentation.  

[Week 3 Assignment](https://github.com/fentontaylor/DevelopingDataProducts/tree/gh-pages/Project2)